### 2017-10-19 memos/generalities.tex
clarify generalies text

\section*{Generalities}

Define the occurrence rate density, $\Gamma$,
as the expected number of planets per star per natural logarithmic bin 
of planetary and stellar phase space:
\begin{equation}
\Gamma(\vec{x}) = \frac{d^n\Lambda}{\prod_{i=1}^{n} {\rm d} \ln x_i  }.
\label{eq:rate_density_defn}
\end{equation}
$\vec{x}$ is an $n$-dimensional list of the continuous physical properties 
that might affect the occurrence rate density. 
For example,
$\vec{x}=(r,P,R)$ where $r$ is the planet radius, $P$ is its orbital period, 
and $R$ is the host star radius.
The occurrence rate $\Lambda$ is found by integrating the rate density over a 
specified volume of phase space.

The previous definition implicitly marginalizes the rate density over stellar 
multiplicity.
For simplicity, this work only considers single and binary star systems.
Then for a selected population of stars and planets the rate density can be 
written
\begin{equation}
\Gamma(\vec{x})
= \sum_{i=0}^{2} w_i \Gamma_i(\vec{x})
= \sum_i w_i \Lambda_i p_i(\vec{x})
\label{eq:rate_density_marginalized}
\end{equation}
where $i=0$ corresponds to single star systems, $i=1$ primaries of binaries, 
and $i=2$ secondaries of binaries.
$\Lambda_i$ is the occurrence rate integrated over all possible phase space  
for the $i^{\rm th}$ system type, $p_i(\vec{x})$ is the joint 
probability density function so that $\Gamma_i(\vec{x}) = \Lambda_i 
p_i(\vec{x})$, 
and the weights are given by 
\begin{equation}
w_i = N_i/N_{\rm tot},
\end{equation}
for $N_{\rm tot} = \sum_i N_i$ the total number of selected stars, and 
$N_0,N_1,N_2$ the number of selected single stars, primaries, and 
secondaries respectively\footnote{
Eq.~\ref{eq:rate_density_marginalized} follows by writing the $i^{\rm 
th}$ system type's rate density as some normalization multiplied by a 
probability density:
$\Gamma_i(\vec{x}) = \mathcal{Z}_i p_i(\vec{x})$.
For Eq.~\ref{eq:rate_density_defn} to hold, we must have $\mathcal{Z}_i = 
\Lambda_i$.
}.
The relationship between the rate $\Lambda$ over a desired volume of 
phase space $\Omega_{\rm desired}$ and $\Lambda_i$ is
\begin{equation}
\Lambda = \sum_i
\left[
\left(
w_i \Lambda_i \int_{\Omega_{\rm desired}} p_i(\vec{x}) \,{\rm d}\Omega
\right)
\cdot
\left(
w_i \Lambda_i \int_{\Omega_{\rm total}} p_i(\vec{x}) \,{\rm d}\Omega
\right)^{-1}
\right],
\label{eq:occ_rate}
\end{equation}
where the inverse term is unity if $p_i(\vec{x})$ is appropriately normalized.

A transit survey will have a rate density of detected planets $\hat{\Gamma}$, 
which will be the (dot) product of the rate density and the detection 
efficiency $Q(\vec{x})$:
\begin{equation}
\hat{\Gamma}(\vec{x}) = \sum_i Q_i(\vec{x}) \Gamma_i(\vec{x}) 
\equiv \sum_i \hat{\Gamma}_i(\vec{x}),
\label{eq:detected_rate_density}
\end{equation}
where again the index $i$ is over each type of system (singles, primaries, 
and secondaries).
The detection efficiency includes the geometric transit probability, as well 
as any incompleteness effects.