### 2017-10-19 paper/main.tex
moving things around

\documentclass[12pt,modern]{aastex61}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{todonotes}

\shortauthors{LGB, KM, JNW}
\shorttitle{Toy transit surveys}

\begin{document}

\title{ Binarity's effects on occurrence rates measured by transit surveys}

\author{
LGB, KM, JNW
}

\input{generalities}
\input{model_1}
\input{model_2}

\end{document}