### 2017-10-20 paper/main.tex

\documentclass[12pt,modern]{aastex61}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{comment}

\shortauthors{LGB, KM, JNW}
\shorttitle{Toy transit surveys}

\begin{document}

\title{ The effects of binarity on planet occurrence rates measured by transit 
surveys}

\author{
L. G. Bouma, K. Masuda, J. N. Winn
}

\begin{abstract}

The aim of this work is to clarify what biases, if any, stellar binarity 
introduces to occurrence rates inferred from transit surveys.
Ignoring binarity leads to diluted planetary radii, overestimated 
detection efficiencies, and an undercounted number of selected stars.
These effects skew occurrence rate measurements in different directions, and 
we develop simplified analytic and numerical models to clarify which are the 
most important.
For models in which all planets have the same radii, we find that 
ignoring binarity leads to an inferred occurrence rate at the true radius two 
to three times smaller than the actual rate (an upper limit to the possible 
error).
%However, the direction and size of the error depend on the assumed 
%planet population's radius distribution.
More realistically, using Howard et al. (2012)'s radius distribution and a 
simple stellar model, we find that ignoring binarity leads to a $\sim10-30\%$ 
underestimate of the number of planets per star over most planetary radii.
In particular, the 30\% underestimate applies to hot Jupiter occurrence rates, 
which suggests that the discrepancy between rates measured by {\it Kepler} and 
RV surveys might be attributable to the {\it Kepler} sample containing more 
binaries.

\end{abstract}

\input{introduction}
\input{methods}
\input{results}
\input{discussion}

\input{figures.tex}

\end{document}