### 2017-10-20 paper/model_3.tex

\subsection{Model \#3: Fixed primaries, varying planets and secondaries}
\label{sec:model_3}

See Fig.~\ref{fig:errcases_model_3_log}.