### 2017-10-28 paper/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

This work presented three simple models for the effects of binarity on 
occurrence rates measured by transit surveys.