### 2017-10-28 paper/conclusion.tex
paper: add intro/disc/conclusion/HJ table

\section{Conclusion}
\label{sec:conclusion}

This work presented three simple models for the effects of binarity on 
occurrence rates measured by transit surveys.
The most realistic of these models (Model \#3) suggests that binarity leads to 
underestimates of occurrence rates at the $\lesssim 30\%$ relative 
error level.
The effects suggest that hot Jupiter rates measured by transit surveys are 
biased to infer $\approx 2$ fewer hot Jupiters per thousand single stars.
Our third model also indicates that binarity's effects on the measured 
occurrence rates of Earth-sized planets are far smaller than current 
systematic uncertainties.
Though our models are simplistic, their agreement with Barclay \& 
Collaborator's recent detailed simulations indicate that they capture the 
important ingredients.