### 2017-10-28 paper/main.tex

\documentclass[12pt,modern]{aastex61}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{comment}

\shortauthors{LGB, JNW}
\shorttitle{Binarity and Occurrence Rates}

\begin{document}

\title{ The effects of binarity on planet occurrence rates measured by transit 
surveys}

\author{
L. G. Bouma, J. N. Winn
}

\begin{abstract}

This work aims to clarify the biases that stellar binarity introduces
to occurrence rates inferred from transit surveys.
In general, stellar multiplicity leads to diluted planetary 
radii, overestimated detection efficiencies, and an undercounted number of 
selected stars (though possibly an overcounted number of searched stars).
These effects skew occurrence rate measurements in different directions, and 
we develop simple models that allow us to understand the crucial effects.
For models in which all stellar systems are either single or twin binaries, 
and all planets are identical, we find that ignoring binarity leads to a
two-to-threefold underestimate of the occurrence at the true radius.
We believe this represents an upper bound on binarity's effects.
Using more realistic models for the stellar population and planetary 
radii, we find that ignoring binarity leads to a $10$\,-\,$30\%$ 
underestimate of the number of planets per star, depending on planet radii.
For the occurrence of Earth-sized planets, in our most realistic model the 
underestimate is by $\approx 15\%$~--~at present far smaller than other 
systematic uncertainties.
For hot Jupiters, we find that the inferred occurrence rate is 
$\approx 1.3\times$ smaller than the true rate around single stars.
We suggest that this latter effect contributes to the discrepant hot Jupiter 
rates measured by {\it Kepler}\ and the California group's respective surveys.
\end{abstract}

\input{introduction}
\input{methods}
\input{results}
\input{discussion}
\input{conclusion}

\acknowledgements
It was a pleasure to share discussions with Kento Masuda, who pointed us in 
this direction, and helped clarify that something like this was worth studying.

\input{figures.tex}
\input{hj_rates.tex}

\end{document}